\chapter{Discord}

Shiva brought a mango-tree log to the front porch, along with the small axe. He set about chopping it into smaller pieces, calling for Rahul as he worked. ``Rahul...''

Rahul came out onto the porch. ``Yes, bhaiya?'' he answered.

``Go and get some fresh mango leaves for the pooja,'' Shiva said.

``I was about to wash the fruits for the prasad,'' Rahul replied.

``Fine. Leave the fruits in water and bring the leaves first.''

``Okay,'' Rahul said, and went back inside.

The log was hard to break. A knot in it made the work stubborn; only small pieces occasionally split away, while most of his blows simply recoiled. The cold did him no favours either, and sitting on the freezing porch made it worse still—the numbness in his hands made each strike harder to bear. He kept at it anyway, as if he had turned it into a contest between himself and the log.

He had grown so absorbed in it that he didn't notice someone laughing at him. He looked up to find Shreya standing there, while Harsh, her elder brother, walked into the veranda. Her mocking smile embarrassed him somehow, though the happiness of her arrival won out. Her hair was done up in such a charming way, matching the colour of her suit, which was laced with golden edges; and the smile on top of it was more than he could bear—harder than the axe's recoil.

He stayed quiet a while, as if time had stopped, before he was jolted back by Harsh.

``Is it too knotted?''

``Yes, it is,'' he replied, as Shreya moved on toward the veranda.

Shiva stood up and was about to call out, but Rahul appeared first.

``Rahul, bring water and something sweet, and ask Mami to prepare tea,'' he said.

``But I was just leaving for the mango leaves you asked for,'' Rahul said, a little frustrated by the ever-changing demands.

``It's okay—let him go for the leaves. I'll manage the rest here. I haven't come as a guest, to enjoy myself and leave,'' Shreya said.

Harsh, meanwhile, noticed Nani, who had just come out into the veranda, and went over to her.

``Sure, why not? May I have a cup of tea too, please?'' Shiva asked with mischief, still unable to pull his eyes away from her.

After a while, Harsh came out to the porch, where Shiva was still wrestling with the log.

``It won't break like this,'' he said.

Shiva turned and asked, ``What should I do, then?''

``Bring a chisel and hammer—they'll do the job cleanly.''

``Wait.'' Shiva dropped the axe and came back with them.

Harsh sat down, held the chisel in one hand and the hammer in the other, and with just a couple of strikes it sank into the wood. The next went in a little farther along, and cracks began to show.

``Thanks—I'll do the rest,'' Shiva said, stepping in.

``Sure.'' Harsh handed him the hammer and walked back to the veranda, to Nani. The two fell into conversation.

Soon Shreya came over with a tray, a cup of tea on it. ``Tell me if it's good,'' she said, drawing his attention.

``Does it matter? If you've made it for me, it'll always be good,'' he said with a smile.

She stepped closer and leaned in, bringing the tray within his reach, and he lifted the cup. He took a sip and gestured to show how good it was.

Meanwhile, Shreya gathered up the split wood and carried it into the veranda, near Nani.

``That's enough, Shiva. Put the rest inside,'' Nani called to him.

After putting everything back, he returned to the veranda, to Nani. Rahul had already brought the fruits over, all washed, and Shiva, along with Rahul and Harsh, began cutting them into pieces. Shreya joined Nani and went inside to prepare the other items.

``Shiva, when is Vinay coming?'' Nana asked, setting his newspaper aside.

Shiva glanced up at the wall clock. ``He should be here any moment now. I told him we'd leave for the temple by nine, and it's already eight-thirty,'' he replied.

Mama came in with a set of carpets, bundled and tied neatly together.

Soon enough, Vinay arrived with his bag of manuscripts. He shook hands with Shiva and greeted Nana and Mama before sitting down beside Mama on the sofa.

``How is bhaiya now?'' Mama asked.

``He's better now than before. The medication's ongoing. I'm taking him to the doctor for another scheduled visit,'' Vinay replied.

In no time, Shreya brought each item out to the veranda, one by one, as Nani prepared them inside. With a practised ease, Nana took the lead, the boys carrying the items behind him. Rahul followed on a bicycle with the carpets. Mama brought out his motorcycle for Nani, and Shreya rode with her brother in the car, which carried the heavier buckets of prasad.

Reaching the temple, they found the barber's wife, Guni, who had already cleaned the place. They began arranging the items—Mama and Harsh spread the carpets, and Vinay prepared the altar, while Shreya and Nani sat on the chabutara under the huge neem tree. Slowly, people began to arrive. The men sat a little apart, some under the tree, some on chairs pulled from a nearby hut. The women sat closer, on the carpets. It was an open temple, with almost no structure to it—just a statue and a pedestal. The Harijan women from the nearby houses arrived and sat to the right, a little apart from the other women. The children gathered at the edges, caught up in their own games.

Nani and Nana sat in front of the altar, and Vinay began his chants. Shreya sat beside them, helping out with whatever item was needed next for the ritual. Shiva went back to the other men near the hut, a few feet away, who were discussing what bad shape the temple was in and how the villagers ought to collect funds to renovate it. Everyone had their own idea—one wanted marble on the walls, another loudspeakers—each pressing the advantages of their own suggestion against what was actually needed, pausing now and then to fold their hands and hail the goddess whenever Vinay blew the conch.

At the end, Vinay blew the fifth conch, announcing the close of the ritual. Shiva called Rahul in and took up the next task—distributing the prasad. He took the bucket of flour prasad, and Rahul took the fruits. Once Rahul had started, Shiva picked up his bucket and began from the end opposite Rahul's. He handed a paper bowl to the first woman in the line and began scooping out the flour with a metal bowl to drop into the one he'd given her. He glanced up to find she hadn't held it up to receive the prasad; he fumbled a little and asked her to hold it out straight, which she did. He put in the flour, then realised the bowl he'd given her was one of the smaller ones, meant for the fruits. He went back to fetch the larger ones, and overheard one woman grumbling to another—``What has he done? How can we take this prasad now? Doesn't he have the sense to start from this side?''—as she slipped into her mouth the piece of apple Rahul had given her.

Now Shiva found himself in a dilemma. Should he go back to the women waiting for their larger bowls, or do what was expected of him? He picked up the bundle of larger bowls and the bucket and went back to the same woman, who was still holding the small one. He walked slowly along the whole line, giving out the prasad one by one, smiling as the older women offered him their blessings. At last, having finished with that group, he turned back to start on the next.

The front group of women—the ones who had been fussing earlier—were looking at him, and then at Nani, with an expression he couldn't read.

Then Nani stepped in. ``Take that bucket of flour near the goddess's pedestal—it has enough in it, or else you'll have to come back midway through the distribution,'' she said, pointing, then smiled at him.

Shiva did as she said, and went on distributing the prasad and receiving the blessings.

By afternoon, they were all back home. Shiva headed to his room, while the elders stayed in the veranda talking and Shreya went inside to Mami. He lay on his bed, facing the wall, his eyes settling on the picture of Lord Rama that hung on it—seated beside an old woman, eating the berries from her hand, his brother at his side.

His peace broke at a knock on his open door. His gaze shifted to it: Shreya stood there with two plates in her hands.

``May I come in?'' she asked, a tease in her voice.

Shiva sat up on the bed. ``Why not? It's already open for you,'' he said with a smile.

She sat down beside him, setting the food on the bed. ``You didn't come with the others for lunch, so I brought it here myself.''

``Thanks. It's better to eat in your company, I guess,'' he said.

``My friends say my company's the best,'' she replied, a touch of attitude in her voice.

``Sure, sure,'' he said, laughing.
She made up the first bite and held it in front of his mouth. He paused. She gestured with her eyes for him to open it, and Shiva went along and took it. Then he waited for her to make up the next.

She looked up at him. ``That was it—now use your own hand.''

``I'm not going to touch it at all now. You started this, so you'll have to finish it,'' he said.