\chapter{Ingratitude}

Evening began to set in. Shiva moved up to the roof, joined by Shreya. The dim sun that had risen through the day was already giving way to the inevitable fog. The nearby orchard, once packed with cranes, was empty now. The season must have driven the birds from their homes—though they always returned. Shiva laid a mat on the cold floor, and the two of them sat together, watching the sun sink slowly below the horizon.

``Isn't it colder here?'' Shiva asked.

``I'm not going to your room—what if someone notices us? And it's a nicer spot anyway, so calm,'' Shreya replied, a little frustration in her voice.

``Okay, okay... but why sit here, by the bulkhead?'' Shiva asked, expecting a reason.

``Remember this wall? We stood right here last year, watching that man,'' she said, smiling. Then her tone shifted abruptly. ``I miss you so much, being so far from you.''

``Hey...'' Shiva lifted her lowered head. ``Why do you worry so much? One day we'll be together for good,'' he continued.

``I don't want to leave. I just want this moment to freeze for us. I could never go down those stairs,'' she said, her voice deepening.

``But then Harsh bhaiya will have to skip his morning class,'' he said with a smile, and she broke into a mix of laughter and tears.

``One day I'll give you everything, and then we won't have to hide. In fact, I've already planned a few places for us to visit together,'' he said.

``Which places? But no heights, please—I'm afraid of falling,'' she said, slowly wiping her tears.

``What about Manali? I've heard it's a really nice place, with lots of snow,'' he said.

``But isn't it on a mountain?'' she pressed.

``Oh, yes. And now that you've put mountains on your enemy list, I don't want to visit them ever either. Besides, what do they even have? Just some rocks with solid water,'' he answered.

``I like the sea,'' she replied.

``Ohh... like in Goa? But people drink there, you know,'' he said, teasing.

``I don't like people who drink. Then we'll go to a forest somewhere,'' she said.

``Who asks to go to a forest? You really are unique,'' Shiva laughed.

``It doesn't matter where we go, as long as I'm with you,'' she said, moving closer and hugging him, laying her head on his chest as if hiding.

``Hey...'' Shiva turned his head down to look at her. He moved aside the strands of hair fallen across her face and gently kissed her forehead. Shreya sat up, out of his embrace. Shiva's eyes locked with hers, and for a moment Shreya got what she wanted—time froze. Shiva moved closer to her. He could feel the warmth of her breath now, despite the cold breeze moving all around them. She leaned in and kissed him. Time didn't just freeze—it disappeared, for them both. Shiva wrapped his arms around her as she sank closer. Moments later she moved lower and rested her head back against his chest.

His eyes fell on his arm wrapped around her—the dark mark, faded though it was, unsettled him still. 