\chapter{Ebb}

The sun shone dimly, giving out its warmth; the afternoon was pleasant. Shiva headed out of class after everyone else had left. Hitching his bag onto his right shoulder, he walked along the second-floor corridor and straight out to the playground. Sitting at the edge of the ground, where the grass grew a little taller, he noticed the volleyball court, its net taken down. The makeshift bamboo poles they'd used for the net stood bare, almost meaningless now. Grass had begun to grow back over the soil that the players' feet had once trodden bare.

Looking up, his gaze fell on Mr. Verma, hurrying along the lane past the ground to reach the class. Shiva got up and jogged over to him.

``Hey, Shiva, how are you?'' Mr. Verma asked.

``I'm good, sir,'' he replied, slowing to match Mr. Verma's pace. ``So, it's maths today, then?'' Shiva added, glancing at the RD Sharma book in Mr. Verma's hand.

``Yeah. I've picked some questions from the Triangles chapter for practice today,'' Mr. Verma replied as they carried on toward the stairs.

``Yeah, I tried some of the questions from this chapter,'' Shiva replied without much enthusiasm.

``Great—so you've already started it!'' Mr. Verma exclaimed.

``Sort of,'' Shiva replied.

``How do you always start so early, in every chapter?'' Mr. Verma asked.

``My Nana is a retired maths lecturer. So since the beginning, we've solved problems together—even ones that don't matter for exams,'' Shiva replied.

``Oh, that explains it. There's a reason I call you the Master Blaster of Maths,'' Mr. Verma laughed.

They went up the stairs and along the corridor, and reached the class. Mr. Verma set his book down in its usual spot. Shiva looked over to find the girls had swapped their seats. He pulled another bench over from the side and sat down, keeping his bag to his left.

``So, in today's class we discussed triangles. I've chosen some questions for practice,'' Mr. Verma said, and began drawing a diagram on the board, the heavy book still in his other hand.

Shiva opened his book and began solving the question, finishing before Mr. Verma had even written it on the board. Then he sat fully engrossed, watching how Mr. Verma went about solving it.

After a few questions, Gauri set her book to her left and slid a little to the right.

``Okay, so that's all for today. Please try to solve the remaining ones I asked you to mark. These will be really important for the exams,'' Mr. Verma said, closing his book and rubbing his hands together to get the chalk dust off them.

``Sir, will you give us a party for the coming New Year?'' Gauri asked.

``Yes, sir. Students from other coaching classes I've met are having one too. Can we please have one as well?'' Roshini joined in, sounding almost pleading, yet warm.

``What sort of party? It'll be really cold by then—look how cold it is already,'' Mr. Verma replied.

``We can just bring a cake and cut it together,'' Roshini replied.

``But I won't be here then. The vacation starts before that, and I'll probably be back by the 5th of January,'' Mr. Verma said.

``It's fine. We can do it on the 5th, then,'' Gauri said.

``Shiva, what do you think? Can we have it?'' Mr. Verma asked, turning to him.

``Shiva, I'll break your nose if you say no,'' Roshini called out with a laugh.

Shiva raised his eyebrows. ``What choice do I have?'' he said, ending in a smile. ``But when do the vacations start?'' he asked. 

``From the 25th,'' Gauri said, turning to him. He nodded.

``And I have to catch the train tomorrow, after school. So technically, this is our last class of the year,'' Mr. Verma added with a satisfied smile.

``So, just three days of school left,'' Gauri pointed out.

``Does it matter? The vacation—if you can even call it that—is only ten days,'' Roshini replied.

``Then the DM's orders will keep coming for holidays anyway, as the weather keeps dipping and rising,'' Shiva replied, zipping up his bag.

``Sorry, everyone, I have to go—my wife's calling. I promised I'd do the packing while talking it over with her, as if I were a child,'' Mr. Verma laughed, bringing the phone to his ear.

Gauri and Roshini blushed at that. Shiva stood up and walked out behind Mr. Verma, his bag back on his right shoulder.

He walked with Mr. Verma to his bicycle near the stand. Unlocking it, he started walking it back toward the central lane that led to the exit gate. The rule against riding your bicycle on the premises had always felt bogus to him, especially when the place was this empty.

Continuing ahead, he saw Gauri and Roshini standing at the crossing, talking. He moved past them, off to one side, without joining their conversation.

``Oh, Mrs. Slim Fit—where are you off to?'' he heard Roshini call from behind.

He paused and turned back to see the girls closing in. ``You're such a bad girl,'' he threw at Roshini.

``By god, that slim waist, cinched in by that tight belt—if you were a girl, I'd have proposed to you right here on the road,'' Roshini broke into laughter.

``What the hell? What rubbish comes into your mind,'' Shiva laughed, almost embarrassed.

``Let me show you,'' Roshini said, looking around. ``There's a good flower,'' she added, and moved off to get one.

``Is she even a girl, or what?'' Shiva said, turning to Gauri.

``Ask her yourself,'' Gauri laughed, covering her face with her handkerchief. ``But first, tell me, where were you? I was waiting for you before the tuition. I'd brought aloo-paratha today—it took me an hour to make,'' she added, her laugh fading into sternness.

``I was busy somewhere,'' he replied flatly.

``Oh, what the hell. It hasn't even been a moment, and you've already betrayed me for someone else,'' Roshini said, returning with a marigold in her hand, her furious face hiding the laughter beneath—making Gauri blush.

Shiva laughed, taking the flower from Roshini. ``Thanks, ma'am. I'll keep it close,'' he replied mockingly, as they reached the gate. Beyond it, he swung onto his bicycle and sped off toward the main road, and was gone.

The vacation was closing in fast—just a day left before it. A few teachers, the ones from far away, had already left. The class after lunch was left free, its teacher one of those who had gone, and Eric sir was assigned to take it. He was an artist himself; he didn't know much Hindi, but he was very playful with the students, even in his fifties. His large belly, straining to rip out of his white shirt, the dark skin peeking through, made for a comical sight. The students loved him for his fun-loving nature. He taught English, mostly to the middle classes.

The class was chattering away, busy gossiping ahead of the coming vacation. Noise was the only possible outcome.

``Keep quiet, class. Open your books and read,'' Eric sir yelled.

``Sir, can you please teach us something?'' one boy yelled before ducking back into the crowd.

``No... do your own work. You must have homework,'' Eric sir replied, visibly searching for the offender.

``Garima, stop talking, please,'' Eric sir said. His threats were as polite as his belly. 

``Sir, she's talking to me. I'm not talking,'' Garima replied, pointing to her benchmate.

``Sir, please teach us the essay on Saint Valentine from the book,'' another offender raised his voice.

``I've already narrated it many times—no more,'' he replied, still unable to find who'd spoken.

``Sir, then the essay on the train, please,'' one more voice echoed in the already noisy room.

``Leave the train and focus on your own engines,'' sir replied. His pronunciation of the words drew the attention of many, who laughed in unison.

Eric sir's gaze fell again on Garima, who was talking to the other girls.

``Why do you keep talking, dear?'' his question aimed at her.

``Sir, it's not me—they keep talking and pulling me in,'' Garima replied, laughter trickling across her face.

``Darling Garima, oil your own engine—don't look at others','' Eric sir said.

The boys broke into laughter, all eyes turning to her. Garima went pale, as if the sky had fallen on her. She tried to move away from her group, sliding along the bench.

``Sir, she doesn't focus on her own engine—she's only interested in others','' a voice came from the crowd, fuelling the laughter among the boys.

``Are you dumb, or what?'' a voice rose above the surrounding noise. Everyone's gaze shifted to Shiva, now on his feet. ``What the hell are you doing? Do you even have any sense?'' he yelled, so loud it could have been heard from the next classroom. The class fell silent, as if they'd seen a ghost.

He flung his book shut, and it skidded to the ground, before he rushed out of the class. He returned a minute later with Vivek sir, the teacher coordinator.

``Sir, please come with me,'' Vivek sir said to Eric sir as he entered the class. ``And the rest of you, keep quiet—the next period starts in a few minutes,'' he told the students, before walking out with Eric sir.

Shiva sat down in his seat, looking gently at Garima, who had dropped her head onto the bench. 