\chapter{Masks}
It was a pleasant evening. Spring brought its own quiet colors to the world. From the outside, the village appeared still and unchanging, yet within it moved with a restlessness that never truly paused. The sun sank slowly behind distant trees and farmlands, spilling a deep red glow across everything it touched.

From the top floor, Shiva watched flocks of crows and egrets crowd the sky, circling as they prepared to settle for the night in the teak plantation nearby. Hundreds—perhaps thousands—arrived in waves, their white feathers stained red by the dying light as they landed on treetops with practiced certainty. He wondered, not for the first time, how they always found their way back. Somehow, they did. Every evening, without fail.

Lost in the sight and his thoughts, Shiva noticed the ringing of his phone resting against the railing only after it had sounded several times. He picked it up quickly.

“Hello, Mami? What happened?”

“Tea is ready,” Mami spoke. “Come down—and bring Shreya with you. She might be in the room on the second floor.”
“Okay, I’m coming,” Shiva replied.

He stepped out, closing the door behind him, and crossed to the room beside the staircase. The door there was half open. He pushed it gently and stepped inside.

“Hey,” a girl sitting on the couch said, setting her phone aside on the table. “How are you?”
“I’m good,” he replied, settling onto the opposite couch. “How about you? You’re Shreya, if I’m not mistaken.”
“So you already know my name,” she said lightly. “It’s only been a couple of hours since I arrived.”
“Mami told me,” he said.
“Oh,” she smiled. “That explains it.”

He hesitated for a moment, then asked, “So… what are you doing here all alone?”
“I was with Didi,” she said. “She went downstairs a few minutes ago and left me here. I didn’t really know what to do, so I picked up my phone—just waiting for someone to come.” She looked at him again. “And then you arrived.”

“Oh,” Shiva said with a soft laugh, “so I saved you.” She laughed along with him, brief and unguarded.

For a moment, neither of them spoke. Shiva sat there, watching her quietly, until he suddenly asked, his tone turning serious, “Are you a watchman?”

Shreya blinked, clearly caught off guard. “What… what does that even mean?” she asked, genuinely confused.

“You have a nice watch,” he said, pointing to her wrist, the seriousness slipping into mischief. “So I thought—you might be a watch-man.”

She stared at him for a second, then laughed. “Oh, then you’re wrong,” she said, amusement spilling into her voice. “That would make me a watch-woman. But you could be one too.” She pointed to the watch on his wrist. Shiva burst into laughter, and Shreya joined him instantly, the room filling with something light and easy.

The moment broke with another ring of Shiva’s phone. He answered it, laughter settling as he spoke. “Yes, Mami. I’m coming—just a moment.” He slipped the phone back into his pocket.

“Sadly, we’ll have to end this very important meeting,” he said, getting to his feet. “The High Court has summoned us both downstairs. Snacks are ready.”

Shreya smiled and stood up as well. Together, they walked out of the room and down the corridor, dimly lit with bulbs and sky turned dark.

After dinner, Shiva found himself back on the top floor once again. The conversations with the guests had left him feeling unnecessary, like someone occupying space meant for others. He hadn’t known any of them before, and he certainly couldn’t talk freely to Shreya either—not with everyone around, watching. In some way, she seemed like him too, present yet slightly apart.

Up here, he felt calmer. From the edge of the terrace, he looked out as far as his eyes could reach, letting the distance settle him. Music hummed softly through his earphones while his attention drifted to a house across the way. Under a bright bulb, a man was making his bed on the rooftop. There were no stairs—only a narrow ladder leaning against the wall.

The man climbed down to fetch a mattress, returned, and spread it carefully, smoothing it from every side. He inspected each corner before leaving again, this time for pillows. When he came back, he placed them neatly, fluffed them, and lay down for a moment—only to sit up almost immediately, as if remembering something important. The light bulb glowed harshly above him, hanging from a bamboo stick tied to one of the columns rising from the house.

With a tired resignation, the man descended the crooked ladder once more.

Shiva watched the small sequence unfold with quiet curiosity, his gaze fixed on the rooftop across the way, as though it were a story revealing itself piece by piece. Then a sudden tap on his shoulder startled him, and he jumped at once, pulled back into the present.

He turned around to find Shreya standing just behind him. Startled, he pulled the earphones out quickly.
“Yeah… hi,” he said, flustered. “What are you doing here?” His words tumbled out as if he had been pulled away from something important.

“Nothing,” she replied lightly. “I got bored downstairs, so I came up for some fresh air.”
“You almost gave me a heart attack,” he said.
“No, no,” she smiled. “You have a long life ahead of you. Just not yet.”

She tilted her head, watching him. “I called you twice, but you were completely lost. Were you talking to your girlfriend?” She paused, her voice playful. “If I interrupted something important, I’m really sorry. I can go stand somewhere else—there’s plenty of space here.”

Her words were teasing, her smile full of mischief, yet there was something gentle about it too—an innocence that made the mockery feel warm rather than sharp.

“No… no,” he said quickly. “I don’t have a girlfriend.” His cheeks warmed as he spoke—barely visible in the darkness, though a few stray bulbs nearby caught the color on parts of his face. “How would a watchman even get a girlfriend?”

She raised an eyebrow, amused. “So you’ve decided to become a watchman from today onward?”
He smiled back at her. “I could think about it—if that’s your command.”

“Command?” she said, a hint of mock seriousness in her voice. “How could I command you? You don’t even know me. I’m not even your friend.” The words tried to sound firm, but her smile betrayed her.

“Aren’t we already friends?” Shiva replied, mischief clear in his tone. “At least I thought we were. Am I mistaken?”

She laughed softly, finally letting go of the attempt at seriousness. “Why not? We can be friends—if you allow it.”

“Are you on Facebook?” he asked, pulling out his phone.
“Yes,” she said. “I sometimes use it on my Didi’s phone. But I do have my own account.”
“Okay, let me look you up,” he said, fingers moving quickly over the keypad. “Is this you?”
“Yes, that’s my account,” she replied, stepping closer to see the small screen.

For a moment, the distance between them narrowed to almost nothing. He lifted his face slightly. Only a small gap separated them now. In the quiet around them, he could hear her breathing, feel the closeness in a way that made the air seem heavier.

“I’ve sent the request,” he said quietly. “Now it’s your turn to accept. I’ve done my part.”

With that, he eased himself back, breaking the closeness, and leaned against the railing behind him. Neither of them spoke. They stood there for a long while, watching the man on the distant rooftop climb down again—this time with a torch in his hand—only to return moments later carrying a jug of water, moving about his small world as if nothing else existed.